---
title: "Data preparation: Factorial taxes - wave 1"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css"
output:
  bookdown::html_document2:
    number_sections: true
    toc: true
    toc_float:
      collapsed: false
      toc_depth: 2
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE,message = FALSE, cache = TRUE,results = "hold")
knitr::opts_knit$set(root.dir = "../")
options(scipen=999)
rm(list=ls())
options(knitr.kable.NA = '')
Sys.setlocale("LC_MESSAGES", 'es_CL.UTF-8')
Sys.setenv(LANG = "es_CL.UTF-8")
```

```{r}
 if (!require("pacman")) install.packages("pacman")  #si falta pacman, instalar
 if (!require("tinytex")) install.packages("tinytex")#si falta tinytex, instalar
 pacman::p_load(dplyr,haven,sjlabelled,tidyverse,sjmisc,stringi) # librerias
# load(file = "input/data/proc/factorial_ola1.RData")
data01 <- sjlabelled::read_spss(path = "input/data/original/factorial_ola1.sav")
load(file = "input/data/proc/vig_dim.RData")
data01 <- data01 %>% filter(consen==1) %>% data.frame()

deck_cod01<- c("FL_191_DO_","FL_193_DO_", #deck01
               "FL_176_DO_","FL_178_DO_", #deck02
               "FL_161_DO_","FL_163_DO_", #deck03
               "FL_146_DO_","FL_148_DO_", #deck04
               "FL_131_DO_","FL_133_DO_", #deck05
               "FL_116_DO_","FL_118_DO_", #deck06
               "FL_101_DO_ ","FL_103_DO_",#deck07
               "FL_86_DO_","FL_88_DO_",   #deck08
               "FL_71_DO_","FL_73_DO_",   #deck09
               "FL_56_DO_","FL_58_DO_"    #deck10
               )
#  for checking and naming decks_vignettes:
decks01<- data01 %>% select(ResponseId,"finished"=Finished,
                            starts_with("p_a_"),
                            starts_with("p_b_"),
                            matches(paste(deck_cod01, collapse="|")))
# rename decks where:
# example: deck_vig11 = Deck 01, Vignette 1
for (i in deck_cod01) {
  names(decks01) <- names(decks01) %>% str_replace_all(pattern = i,replacement = "deck_vig") 
}

decks01<- remove_label(decks01)

# haven::write_dta(decks01,path = 'input/data/proc/decks.dta')
```

duplicated cases
```{r}
frq1 <- as.data.frame(frq(data01$ticket))

tickets<- data01 %>% 
  #check those cases with more than 1 cases (duplicates)
  filter(ticket %in% filter(frq1,frq>1)$val) %>% 
  select(ticket,ResponseId,RecordedDate,StartDate,EndDate,Progress) %>%
  #keep cases with more than 50% of the survey
  filter(Progress>50) %>% 
  arrange(ticket) %>% 
  data.frame()
dim(tickets)
frq2 <-
    #check if after the filter there are still cases with more than 1
data.frame(frq(tickets$ticket)) %>% 
    #keep those cases that are still duplicated
  filter(frq>1)

tickets<- data01 %>% 
  #keep the duplicated cases and check them
  filter(ticket %in% frq2$val) %>% 
  select(ticket,ResponseId,RecordedDate,StartDate,EndDate,Progress) %>% 
  filter(Progress>50) %>% 
  arrange(ticket) %>% 
  data.table::as.data.table()
dim(tickets)

progress <- 
tickets %>% group_by(ticket) %>% summarise(pmin=min(Progress,na.rm =T),
                                           pmax=max(Progress,na.rm =T),
                                           start_min=min(StartDate,na.rm =T),
                                           start_max=max(StartDate,na.rm =T),
                                           ) %>% arrange(pmax,pmin)

#check those cases with the lower RecordedDate (earliest)
tickets[tickets[, .I[which.min(RecordedDate)], by=ticket]$V1] 

# these are tha cases that we want to delete from the data
delete_ids<- as.character(tickets[tickets[, .I[which.min(RecordedDate)], by=ticket]$V1]$ResponseId )

# Filter the data-----------------------------------------------------------
dim(data01)
data01_f <- 
data01 %>% 
  filter(Progress>50)
table(data01$ResponseId %in% delete_ids)
df1<- data.frame(frq(data01_f$ticket))
data01_f <- 
data01_f %>% 
  filter(!ResponseId %in% delete_ids)
dim(data01_f)
data01 <- data01_f

df01 <- data.frame(frq(data01$ticket))
```



> The median length of the questionnaire is **`r colorize(paste0(round(median(data01$Duration__in_seconds_/60),2)," minutos"),"red")` **  


```{r}
tiempo<- data01 %>% filter(Finished==1,Progress==100,!is.na(Q5231)) %>%
  select(Duration__in_seconds_) %>% summarise(tiempo=median(Duration__in_seconds_/60))
```

> Excluding incomplete answers (finished= 0) the median length of the questionnaire is **`r colorize(paste0(round(tiempo,2)," minutos"),"red")` **  


* We select the variables that identify the vignette and the deck to which it belongs.

```{r}
nam01<- decks01 %>% select(starts_with("deck_vig")) %>% names()
decks02 <- data.frame(!is.na(decks01[nam01]))
decks02$ResponseId <- decks01$ResponseId # add the responseID for merge
table(is.na(decks02$ResponseId)) # no missing in ResponseID
decks02[decks02==FALSE] <- NA # recorde FALSE as NA
names(decks02)

decks02 <- decks02 %>% filter(ResponseId %in% data01$ResponseId)

# select deck01 with responsed for perceived and just income tax
decks03 <- left_join(select(decks01,-nam01),decks02,by="ResponseId")
names(decks03)

# if deck03 is TRUE, inpute column name
w <- which(decks03=="TRUE",arr.ind=TRUE) #the position of the coordinates
decks03[w] <- names(decks03)[w[,"col"]]
```

- Database from wide to long

```{r}
library(tidyr)
# Step 1: wide to long for perceived tax
  # get the variable names for perception and each deck/vignette
vars_perc<- decks03 %>% select(starts_with('p_a_')) %>% names()
  # subset with each case and perception from 1 to 120 (number of vignettes)
perc<- decks03 %>% select(ResponseId,starts_with('p_a_'))

#pivot the data frame into a long format
long_perc <- 
perc %>% pivot_longer(cols=vars_perc,
                    names_to='decks_vig_perc',
                    values_to='perc') %>% 
  arrange(decks_vig_perc)
table(long_perc$perc=="") # FALSE = vignettes with no responses by desing
long_perc$perc[long_perc$perc==""] <- NA #set responses in blank to NA

long_perc$decks_vig_a <- 
  str_replace_all(long_perc$decks_vig_perc,pattern = "p_a_",replacement = "")

# create an common ID
long_perc$ID <- paste0(long_perc$ResponseId,long_perc$decks_vig_a)

# Step 2: wide to long for just tax
  # get the variable names for just and each deck/vignette
vars_just<- decks03 %>% select(starts_with('p_b_')) %>% names()
  # subset with each case and just from 1 to 120 (number of vignettes)
just<- decks03 %>% select(ResponseId,starts_with('p_b_'))

#pivot the data frame into a long format
long_just <- 
just %>% pivot_longer(cols=vars_just,
                    names_to='decks_vig_just',
                    values_to='just') %>% 
  arrange(decks_vig_just)
table(long_just$just=="") # FALSE = vignettes with no responses by desing
long_just$just[long_just$just==""] <- NA #set responses in blank to NA

long_just$decks_vig_b <- 
  str_replace_all(long_just$decks_vig_just,pattern = "p_b_",replacement = "")

# create an common ID
long_just$ID <- paste0(long_just$ResponseId,long_just$decks_vig_b)


# Step 3: merge of perceived and just income tax
table(long_just$ID==long_perc$ID) #all *must* be true 

long <- left_join(x = long_perc,y = long_just,"ID") %>% 
  select(ResponseId=ResponseId.x,
         deck_vig=decks_vig_b,
         taxperc=perc,
         taxjust=just)
long01<- na.omit(long)
```

- Recode perceived and fair tax variables

```{r}
long01$taxperc <- stri_replace_all(long01$taxperc, "", fixed=c("$")) #delete $ symbol
long01$taxperc <- stri_replace_all(long01$taxperc, "", fixed=c(".")) #delete . 
long01$taxperc <- as.numeric(long01$taxperc) # transform to numeric
summary(long01$taxperc)
# sjmisc::frq(long01$taxperc)
# sjPlot::plot_frq(long01$taxperc,type = 'density')

long01$taxjust <- stri_replace_all(long01$taxjust, "", fixed=c("$")) #delete $ symbol
long01$taxjust <- stri_replace_all(long01$taxjust, "", fixed=c(".")) #delete . 
long01$taxjust <- as.numeric(long01$taxjust) # transform to numeric
summary(long01$taxjust)
# sjmisc::frq(long01$taxjust)
# sjPlot::plot_frq(long01$taxjust,type = 'density')
```


-  Merge respondents and vignettes

```{r}
# A) Merge sociodemographics from respondent to long base.
long02<- 
  left_join(long01,data01[,c("ticket","ResponseId","Finished","sexo","edad",
                             "educ","ingresos",'cant_hogar',"act_prin","comuna")])  

#select ID from ticket variable
long02$folio_encuestado_1 <- stringr::str_split(long02$ticket,pattern = "_",simplify = T)[,1]
long02$folio_encuestado_4 <- stringr::str_split(long02$ticket,pattern = "_",simplify = T)[,4]

# B) Name respondent level variables
long02<- 
  long02 %>% 
  rename("respondeid" = ResponseId, "sexo_re" = sexo, "edad_re" = edad, 
         "educ_re" =educ, "ingresos_re" = ingresos,
         pophome="cant_hogar",  
         "estlab_re" = act_prin, 
         "finished" = Finished) 

# C) Naming vignette-level variables (generated in SAS)

vig_dat<- vig_dat %>% rename("sexo_vig"=sexo,
                             "apellido_vig"=apellido,
                             "edad_vig"=edad,
                             "nse_ori_vig"="nivel_socioeconómico_de_origen",
                             "educ_vig"=nivel_educativo,
                             "necesidades_vig"=necesidades,
                             "educ_madre_vig"="educación_de_la_madre", 
                             "ingresos_vig"=ingresos)


# D) Merge bullet data to long base (respondent level + vignettes)
vig_resp<- 
  left_join(long02,vig_dat,by =c("deck_vig"))  %>%
  select(starts_with("folio_encuestado"),respondeid,deck,id,deck_vig,everything(),-p1)  
```

- Variable label 

```{r}
vig_resp <- vig_resp %>% var_labels(respondeid = "Id Respondent",
                                    finished="Finished survey",
                                    deck = "Id Deck",
                                    id = "Id Vignette",
                                    deck_vig = "Id deck and vignette",
                                    taxperc = "Perceived income tax",
                                    taxjust = "Just income tax",
                                    sexo_re = "Gender (respondent level)",
                                    edad_re = "Age (respondent level)",
                                    educ_re = "Education (respondent level)",
                                    ingresos_re = "Income in CLP (respondent level)",
                                    estlab_re = "Job status (respondent level)",
                                    comuna = "Municipality (respondent level)",
                                    sexo_vig = "Gender (vignette level)",
                                    apellido_vig = "Etnicity - surname (vignette level)",
                                    edad_vig = "Age (vignette level)",
                                    nse_ori_vig = "SES of origin (vignette level)",
                                    educ_vig = "Education (vignette level)",
                                    necesidades_vig = "Needs - number of children (vignette level)",
                                    educ_madre_vig = "Mother's education level (vignette level)",
                                    ingresos_vig = "Income in CLP (vignette level)")

```

- Value labels 

```{r vignette}
vig_resp$sexo_vig <- 
  set_labels(vig_resp$sexo_vig,labels = c("Male"=1,"Female"=2))
vig_resp$apellido_vig <- 
  set_labels(vig_resp$apellido_vig,
             labels = c("Spanish"=1,"Basque-Castilian"=2,"European"=3,"Indigenous"=4))
vig_resp$edad_vig <-
  set_labels(vig_resp$edad_vig,
             labels = c("35 years"=1,"45 years"=2,"55 years"=3))
vig_resp$nse_ori_vig<- 
  set_labels(vig_resp$nse_ori_vig,labels = c("Public"=1,"Charter"=2,"Private"=3))
vig_resp$educ_vig <-
  set_labels(vig_resp$educ_vig,
             labels = c("Complete basic education"=1,
                        "Complete secondary education"=2,
                        "Technical professional education"=3,
                        "Full university education"=4,
                        "Graduate Studies"=5))
vig_resp$necesidades_vig<- 
  set_labels(vig_resp$necesidades_vig,labels = c(
    "Do not have children" = 1,
    "Has 1 child" = 2,
    "Has 2 children" = 3,
    "Has 5 children" = 4))
vig_resp$educ_madre_vig <- 
  set_labels(vig_resp$educ_madre_vig ,labels = c(
    "Complete basic education" = 1,
    "Complete secondary education" = 2,
    "Technical professional education" = 3,
    "Full university education" = 4,
    "Graduate Studies" = 5
  ))
vig_resp$ingresos_vig <- 
  set_labels(vig_resp$ingresos_vig,
             labels = c("$320.500"=1,"$655.000"=2,"$1.200.000"=3,
                        "$2.500.000"=4,"$4.300.000"=5))


vig_resp$sexo_vig <- to_factor(vig_resp$sexo_vig, labels = c("Male" = 1, "Female" = 2))
vig_resp$apellido_vig <- to_factor(vig_resp$apellido_vig, labels = c("Spanish" = 1, "Basque-Castilian" = 2, "European" = 3, "Indigenous" = 4))
vig_resp$edad_vig <- to_factor(vig_resp$edad_vig, labels = c("35 years" = 1, "45 years" = 2, "55 years" = 3))
vig_resp$nse_ori_vig <- to_factor(vig_resp$nse_ori_vig, labels = c("Public" = 1, "Charter" = 2, "Private" = 3))
vig_resp$educ_vig <- to_factor(vig_resp$educ_vig, labels = c("Complete basic education" = 1, "Complete secondary education" = 2, "Technical professional education" = 3, "Full university education" = 4, "Graduate Studies" = 5))
vig_resp$necesidades_vig <- to_factor(vig_resp$necesidades_vig, labels = c("Do not have children" = 1, "Has 1 child" = 2, "Has 2 children" = 3, "Has 5 children" = 4))
vig_resp$educ_madre_vig <- to_factor(vig_resp$educ_madre_vig, labels = c("Complete basic education" = 1, "Complete secondary education" = 2, "Technical professional education" = 3, "Full university education" = 4, "Graduate Studies" = 5))
vig_resp$ingresos_vig <- to_factor(vig_resp$ingresos_vig, labels = c("$320,500" = 1, "$655,000" = 2, "$1,200,000" = 3, "$2,500,000" = 4, "$4,300,000" = 5))
```

```{r respondent}
vig_resp$sexo_re <- set_labels(x = vig_resp$sexo_re, 
                               labels = c("Man"=1,"Woman"=2))
vig_resp$sexo_re <- factor(vig_resp$sexo_re,levels = c(1,2),labels = c("Male","Female"))



vig_resp$educ_re <- set_labels(x = vig_resp$educ_re,labels = 
                                 c("No formal education"=1,
                                  "Primary education incomplete"=2,
                                  "Complete primary education"=3,
                                  "Incomplete secondary education"=4,
                                  "Complete secondary education"=5,
                                  "Non-university higher education incomplete"=6,
                                  "Complete non-university higher education"=7,
                                  "Incomplete college education"=8,
                                  "Complete college education"=9,
                                  "Postgraduate studies, master's, doctoral"=10)) 

vig_resp$estlab_re <-
  set_labels(vig_resp$estlab_re,labels = c(
  "Working for pay full-time" = 1,
  "Working for pay part-time or do odd jobs" = 2,
  "Study and work" = 3,
  "Just study" = 4,
  "Retired or pensioned" = 5,
  "Unemployed, looking for work" = 6,
  "Performs unpaid work (household chores, babysitting and others)" = 7,
  "It is ill or has a disability" = 8,
  "Not studying, not working and not looking for work" = 9))


vig_resp$ingresos_re <- 
  set_labels(vig_resp$ingresos_re,
             labels = 
               c("Less than $35.000 monthly disposable income" = 1 ,
                 "From $35.001  to $56.000 monthly disposable income" = 2 ,
                 "From $56.001  to $78.000 monthly disposable income" = 3 ,
                 "From $78.001  to $101.000 monthly disposable income" = 4 ,
                 "From $101.001 to $134.000 monthly disposable income" = 5 ,
                 "From $134.001 to $179.000 monthly disposable income" = 6 ,
                 "From $179.001 to $224.000 monthly disposable income" = 7 ,
                 "From $224.001 to $291.000 monthly disposable income" = 8 ,
                 "From $291.001 to $358.000 monthly disposable income" = 9 ,
                 "From $358.001 to $448.000 monthly disposable income" = 10,
                 "From $448.001 to $1.000.000 monthly disposable incomes" = 11,
                 "From $1.000.001 to $2.000.000 monthly disposable income" = 12,
                 "From $2.000.001 to $3.000.000 monthly disposable income" = 13,
                 "More than $3.000.000 monthly disposable income" = 14))
```

- Recode variables

```{r}
# Household income
sjmisc::frq(vig_resp$pophome)
sjmisc::frq(vig_resp$ingresos_re)
vig_resp$hhincome <-
  car::recode(vig_resp$ingresos_re,
              recodes ="1=35000;2=45500.5;3=67000.5;4=89500.5;5=117500.5;
              6=156500.5;7=201500.5;8=257500.5;9=324500.5;10=403000.5;
              11=724000.5;12=1500001;13=2500001;14=3000000")
vig_resp$hhincome <- as.numeric(vig_resp$hhincome)

sjmisc::frq(vig_resp$hhincome)
vig_resp$income5 <- ntile(vig_resp$hhincome,n = 5)
sjmisc::frq(vig_resp$income5)
vig_resp$income5[is.na(vig_resp$income5)] <- 6
vig_resp$income5 <- factor(vig_resp$income5,levels = c(1,2,3,4,5,6),
                           labels = c("Quintile 1","Quintile 2","Quintile 3",
                                      "Quintile 4","Quintile 5","No income/Missing"))
set_label(vig_resp$income5) <- "Household income quintile (respondent level)"
sjmisc::frq(vig_resp$income5)
# Educational levele

sjmisc::frq(vig_resp$educ_re) 
vig_resp$educ <- car::recode(vig_resp$educ_re,
                             recodes = "c(1,2,3)=1;4:5=2;6:7=3;c(8,9,10)=4",as.factor = T,
                             levels = c(1,2,3,4))

vig_resp$educ <- factor(x =vig_resp$educ ,
                        labels =  c("Primary or less","High school",
                                    "Technical","Universitary"))

set_label(vig_resp$educ) <- get_label(vig_resp$educ_re) 
sjmisc::frq(vig_resp$educ) 
```

-save database

```{r}
vig_respw01 <-vig_resp 
view_df(vig_respw01,show.type = T)
skimr::skim(vig_respw01)
dim(vig_respw01)

vig_respw01<- 
  vig_respw01[which(vig_respw01$folio_encuestado_1!=""),]

save(vig_respw01,file = "input/data/proc/vig_respw01.RData")
```

